---
title: "Descriptive Statistics with **agricolae**"
author: "Felipe de Mendiburu^1^, Muhammad Yaseen^2^"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
  bookdown::pdf_document2:
    dev: cairo_pdf
    fig_caption: yes
    number_sections: yes
    toc: yes
geometry: margin=3cm
documentclass: article
header-includes:
- \usepackage{fancyhdr}
- \usepackage{wrapfig}
- \usepackage{float}
- \pagestyle{fancy}
- \fancyfoot[C]{\thepage}
- \usepackage{hyperref}
- \hypersetup{colorlinks=true}
- \hypersetup{linktoc=all}
- \hypersetup{linkcolor=blue}
- \usepackage{pdflscape}
- \usepackage{booktabs}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \renewcommand\thesection{\arabic{section}}
link-citations: yes
csl: frontiers.csl
resource_files:
- vignettes/rbase.png
- vignettes/rstudio.png
- vignettes/rstudio panes.png
bibliography: Bibliography.bib
vignette: |
  %\VignetteIndexEntry{DescriptiveStats}
  %\usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown_notangle}
---


```{r  echo=FALSE}
out_type <- knitr::opts_knit$get("rmarkdown.pandoc.to")

r = getOption("repos")
r["CRAN"] = "https://cran.rstudio.com/"
#r["CRAN"] = "https://cloud.r-project.org/"
#r["CRAN"] = "https://ftp.iitm.ac.in/cran/"
options(repos = r)
```

```{r results='asis', echo=FALSE}
switch(out_type,
    html = {cat("<p>1. Professor of the Academic Department of Statistics and Informatics of the Faculty of Economics and Planning.National University Agraria La Molina-PERU.</p>
    
<p>2. Department of Mathematics and Statistics, University of Agriculture Faisalabad, Pakistan.</p>")},
    latex = cat("\\begin{center}
1. Professor of the Academic Department of Statistics and Informatics of the Faculty of Economics and Planning.National University Agraria La Molina-PERU.

2. Department of Mathematics and Statistics, University of Agriculture Faisalabad, Pakistan.

\\end{center}" )
)
```


```{r include = FALSE}
knitr::opts_chunk$set(
    echo    = TRUE
  , comment = ""
  , fig.cap = ""
  )
library(agricolae)
```


\begin{center}
\vspace{6pt}
\hrule
\end{center}


# Descriptive statistics
The package **agricolae**  provides some complementary functions to the **R**  program, specifically for the management of the histogram and function *hist*.

## Histogram
The histogram is constructed with the function *graph.freq* and is associated to other functions: *polygon.freq, table.freq, stat.freq*. See Figures: \@ref(fig:DescriptStats2), \@ref(fig:DescriptStats6) and \@ref(fig:DescriptStats7)  for more details. 

Example.  Data generated in **R** . (students' weight).

```{r DescriptStats1}
weight<-c( 68, 53, 69.5, 55, 71, 63, 76.5, 65.5, 69, 75, 76, 57, 70.5, 71.5, 56, 81.5,
           69, 59, 67.5, 61, 68, 59.5, 56.5, 73, 61, 72.5, 71.5, 59.5, 74.5, 63)
print(summary(weight)) 
```


```{r DescriptStats2, fig=TRUE, width=6, height=2, fig.cap="Absolute and relative frequency with polygon"}
oldpar<-par(mfrow=c(1,2),mar=c(4,4,0,1),cex=0.6)
h1<- graph.freq(weight,col=colors()[84],frequency=1,las=2,density=20,ylim=c(0,12),ylab="Frequency") 
x<-h1$breaks
h2<- plot(h1, frequency =2, axes= FALSE,ylim=c(0,0.4),xlab="weight",ylab="Relative (%)")
polygon.freq(h2, col=colors()[84], lwd=2, frequency =2)
axis(1,x,cex=0.6,las=2)
y<-seq(0,0.4,0.1)
axis(2, y,y*100,cex=0.6,las=1) 
par(oldpar)
```


## Statistics and Frequency tables

Statistics: mean, median, mode and standard deviation of the grouped data.
```{r DescriptStats3}
stat.freq(h1)
```

Frequency tables: Use *table.freq*, *stat.freq* and *summary*

The table.freq is equal to summary()

Limits class: **Lower and  Upper**

Class point: **Main**

Frequency: **Frequency**

Percentage frequency: **Percentage**

Cumulative frequency: **CF**

Cumulative percentage frequency: **CPF**

```{r DescriptStats4}
print(summary(h1),row.names=FALSE)
```

## Histogram manipulation functions

You can extract information from a histogram such as class intervals *intervals.freq*, attract new intervals with the *sturges.freq* function or to join classes with *join.freq* function. It is also possible to reproduce the graph with the same creator *graph.freq* or function *plot* and overlay normal function with *normal.freq* be it a histogram in absolute scale, relative or density . The following examples illustrates these properties.


```{r DescriptStats5}
sturges.freq(weight)
intervals.freq(h1)
join.freq(h1,1:3) -> h3
print(summary(h3))
```

```{r DescriptStats6,fig=TRUE  ,width=6, height=2, fig.cap="Join frequency and relative frequency with normal and Ogive"}
oldpar<-par(mfrow=c(1,2),mar=c(4,4,0,1),cex=0.8)
plot(h3, frequency=2,col=colors()[84],ylim=c(0,0.6),axes=FALSE,xlab="weight",ylab="%",border=0)
y<-seq(0,0.6,0.2)
axis(2,y,y*100,las=2)
axis(1,h3$breaks)
normal.freq(h3,frequency=2,col=colors()[90])
ogive.freq(h3,col=colors()[84],xlab="weight")
par(oldpar)
```


## hist() and graph.freq() based on grouped data

The *hist* and *graph.freq* have the same characteristics, only f2 allows build histogram from grouped data.

::: {.verbatim data-latex=""}
 0-10 (3)
 
10-20 (8)

20-30 (15)

30-40 (18)

40-50 (6)
:::


```{r DescriptStats7,fig=TRUE,  width=6, height=1.7, fig.cap="hist() function and histogram defined class"}
oldpar<-par(mfrow=c(1,2),mar=c(4,3,2,1),cex=0.6)
h4<-hist(weight,xlab="Classes (h4)")
table.freq(h4)
# this is possible
# hh<-graph.freq(h4,plot=FALSE)
# summary(hh)
# new class
classes <- c(0, 10, 20, 30, 40, 50)
freq <- c(3, 8, 15, 18, 6)
h5 <- graph.freq(classes,counts=freq, xlab="Classes (h5)",main="Histogram grouped data")
par(oldpar)
```



```{r DescriptStats8}
print(summary(h5),row.names=FALSE)
```



# References
