---
title: "Introduction to **agricolae**"
author: "Felipe de Mendiburu^1^, Muhammad Yaseen^2^"
date: "`r Sys.Date()`"
geometry: margin=3cm
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  pdf_document:
    dev: cairo_pdf
    fig_caption: yes
    number_sections: yes
    toc: yes  
documentclass: article
header-includes:
- \usepackage{fancyhdr}
- \usepackage{wrapfig}
- \usepackage{float}
- \pagestyle{fancy}
- \fancyfoot[C]{\thepage}
- \usepackage{hyperref}
- \hypersetup{colorlinks=true}
- \hypersetup{linktoc=all}
- \hypersetup{linkcolor=blue}
- \usepackage{pdflscape}
- \usepackage{booktabs}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \renewcommand\thesection{\arabic{section}}
link-citations: yes
csl: frontiers.csl
resource_files:
- vignettes/rbase.png
- vignettes/rstudio.png
- vignettes/rstudio panes.png
bibliography: Bibliography.bib
vignette: |
  %\VignetteIndexEntry{Introduction}
  %\usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown_notangle}
---


```{r  echo=FALSE}
out_type <- knitr::opts_knit$get("rmarkdown.pandoc.to")

r = getOption("repos")
r["CRAN"] = "https://cran.rstudio.com/"
#r["CRAN"] = "https://cloud.r-project.org/"
#r["CRAN"] = "https://ftp.iitm.ac.in/cran/"
options(repos = r)
```

```{r results='asis', echo=FALSE}
switch(out_type,
    html = {cat("<p>1. Professor of the Academic Department of Statistics and Informatics of the Faculty of Economics and Planning.National University Agraria La Molina-PERU.</p>
    
<p>2. Department of Mathematics and Statistics, University of Agriculture Faisalabad, Pakistan.</p>")},
    latex = cat("\\begin{center}
1. Professor of the Academic Department of Statistics and Informatics of the Faculty of Economics and Planning.National University Agraria La Molina-PERU.

2. Department of Mathematics and Statistics, University of Agriculture Faisalabad, Pakistan.

\\end{center}" )
)
```


```{r include = FALSE}
knitr::opts_chunk$set(
    echo    = TRUE
  , comment = ""
  , fig.cap = ""
  )
library(agricolae)
```


\begin{center}
\vspace{6pt}
\hrule
\end{center}



# Introduction

The package **agricolae** offers a broad functionality in the design of experiments, especially for experiments in agriculture and improvements of plants, which can also be used for other purposes. It contains the following designs: lattice, alpha, cyclic, balanced incomplete block designs, complete randomized blocks, Latin, Graeco-Latin, augmented block designs, split plot and strip plot. It also has several procedures of experimental data analysis, such as the comparisons of treatments of Waller-Duncan, Bonferroni, Duncan, Student-Newman-Keuls, Scheffe, Ryan, Einot and Gabriel and Welsch multiple range test or the classic LSD and Tukey; and non-parametric comparisons, such as Kruskal-Wallis, Friedman, Durbin, Median and Waerden, stability analysis, and other procedures applied in genetics, as well as procedures in biodiversity and descriptive statistics, @Mend2009


## Installation
The main program of **R** should be already installed in the platform of your computer *(Windows, Linux or MAC)*. If it is not installed yet, you can download it from the R project [https://www.r-project.org/](https://www.r-project.org/) of a repository CRAN.


```{r eval=FALSE}
install.packages("agricolae")
```

Once the `agricolae` package is installed, it needs to be made accessible to the current **R**   session by the command:


```{r results = "hide"}
library(agricolae)
```

For online help facilities or the details of a particular command
(such as the function `waller.test`) you can type:

```{r eval=FALSE}
help(package="agricolae")
help(waller.test)
```


For a complete functionality, `agricolae` requires other packages

<!-- --> | <!-- --> 
---------|---------------------------------------------
[`MASS`](https://cran.r-project.org/web/packages/MASS/index.html):|	for the generalized inverse used in the function `PBIB.test`
[`nlme`](https://cran.r-project.org/web/packages/nlme/index.html): | for the methods REML and LM in `PBIB.test`
[`klaR`](https://cran.r-project.org/web/packages/klaR/index.html):  | for the function `triplot` used in the function `AMMI`
[`cluster`](https://cran.r-project.org/web/packages/cluster/index.html):  | for the use of the function `consensus`
[`AlgDesign`](https://cran.r-project.org/web/packages/AlgDesign/index.html): | for the balanced incomplete block design `design.bib`


## Use in **R**

Since **agricolae **  is a package of functions, these are operational when they are called directly from the console of **R**  and are integrated to all the base functions of **R**.
The following orders are frequent: 

```{r }
detach(package:agricolae) # detach package agricole
library(agricolae) # Load the package to the memory
designs<-apropos("design")
print(designs[substr(designs,1,6)=="design"], row.names=FALSE)
```

::: {.verbatim data-latex=""}
For the use of symbols that do not appear in the keyboard in Spanish, such as:

~, [, ], &, ^, |. <, >, {, }, \% or others, use the table ASCII code.
:::



```{r}
library(agricolae) # Load the package to the memory: 
```

In order to continue with the command line, do not forget to close the open windows with any R order.
For help: 
```{r eval=FALSE}
help(graph.freq)
? (graph.freq)
str(normal.freq)
example(join.freq)
```


## Data set in **agricolae **
```{r}
A<-as.data.frame(data(package="agricolae")$results[,3:4])
A[,2]<-paste(substr(A[,2],1,35),"..",sep=".")
head(A)
```

# References
